---
title: "Question7,8"
author: "Thisha Thiagarajan A15474979"
date: "12/2/2021"
output: html_document
---
Read in the fasta formatted alignment. 
```{r}
library(bio3d)
alignment <- read.fasta("sample.fasta", rm.dup = TRUE, to.upper = FALSE, to.dash=TRUE)
```

Calculate the sequence identity matrix

```{r}
heatmap_data <- seqidentity(alignment, normalize=TRUE, similarity=FALSE, ncore=1, nseg.scale=1)
```

Create heatmap plot 

```{r}
heatmap(heatmap_data, margins =c(12,12))
```
Generate a consensus sequence from your alignment

```{r}
consensus(alignment)
```
Consensus alignment has a lot of gaps. I will use an original sequence from the alignment for my search of the PDB. I chose the sequence with the highest identity to all others in your alignment by calculating the row-wise maximum from my sequence identity matrix.

```{r}
apply(heatmap_data[1:10,], 1, sum)
```

Using this analysis, I will use the sequence from HomoSapien_APOEQuery. 

```{r}
pdb <- read.pdb("6V7M")
blast <- blast.pdb(pdbseq(pdb) )

```

```{r}
blast
# 
```

