---
title: "class17"
author: "Thisha Thiagarajan A15474979"
date: "12/3/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
## Understanding the Data 

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
tail(vax)
```
> Q1. What column details the total number of people fully vaccinated? 

vax$persons_fully_vaccinated

> Q2. What column details the Zip code tabulation area? 

vax$zip_code_tabulation_area

> Q3. What is the earliest date in this dataset?

2021-01-05

> Q4. What is the latest date in this dataset?

2021-11-30

```{r}
skimr::skim(vax)
```

> Q5. How many numeric columns are in this dataset? 

9

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
sum( is.na(vax$persons_fully_vaccinated) )
```


8472

> Q7. What percent of persons_fully_vaccinated values are missing?

10.0%

```{r}
8472/84672 
1-0.900
```

> Q8. [Optional]: Why might this data be missing?

This data may be missing because this surverying tools used to gather this data did not get enough responses. Or the researchers may not have access to that data. 

## Working with Dates 

```{r}
library(lubridate)
today()
vax$as_of_date <- ymd(vax$as_of_date)
```
```{r}
today() - vax$as_of_date[1]
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]

```
> Q9. How many days have passed since the last update of the dataset?

3 days

```{r}
today() - vax$as_of_date[84672]

```

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

There are 48. 

```{r}
#unique_count = n_distinct(vax$as_of_date)
#unique_count
unique_count = length(unique(vax$as_of_date))
unique_count
```

## Working with ZIP codes

```{r}
#install.packages("zipcodeR")
library(zipcodeR)
geocode_zip('92037')
zip_distance('92037','92109')
reverse_zipcode(c('92037', "92109") )

#To Pull data for all ZIP codes in the dataset
#zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

## Focus on the San Diego area

```{r}
sd <- vax[vax$county == "San Diego", ]
nrow(sd)
```

Can also do this with dpylr

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
head(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

> Q11. How many distinct zip codes are listed for San Diego County?

There are 107 distinct zip codes. 

```{r}
#unique_zipcode = n_distinct(sd$zip_code_tabulation_area)
#unique_zipcode

#can also do this with unique and length
test_unique = length(unique(sd$zip_code_tabulation_area))
test_unique
```

> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 

92154

```{r}
row_largest12 <- sd[which.max(sd$age12_plus_population),]
row_largest12$zip_code_tabulation_area
```

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?

The overall average is 0.6722183. 
67.2%

```{r}
sd.date <- filter(vax, county == "San Diego" &
                as_of_date == "2021-11-16")
mean(sd.date$percent_of_population_fully_vaccinated, na.rm = TRUE)
```

> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
hist(sd.date$percent_of_population_fully_vaccinated, main = "Histogram of Vaccination Rates Across San Diego County", xlab = "Percent of Population Fully Vaccinated on 2021-11-09", ylab = "Frequency")
```

## Focus on UCSD/La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
head(ucsd)
ucsd[1,]$age5_plus_population
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
library(ggplot2)

ggplot(ucsd) +
  aes(x = ucsd$as_of_date,
      y = ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```

## Comparing 92037 to other similar sized areas?

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```

> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

The calculated mean is 0.6645132. 

```{r}
mean_vax36 <- mean(vax.36$percent_of_population_fully_vaccinated)
mean_vax36
```
```{r}
ggplot(ucsd) +
  aes(x = ucsd$as_of_date,
      y = ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  geom_hline(yintercept = 0.6629812, linetype= "dashed") +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

Min: 0.353269
1st Quartile: 0.591029
Median: 0.666919
3rd Quartile: 0.731112
Max: 1.000000
Mean: 0.6645132

```{r}
fivenum(vax.36$percent_of_population_fully_vaccinated)
mean(vax.36$percent_of_population_fully_vaccinated)
```

> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) +
  aes(x = percent_of_population_fully_vaccinated) +
  geom_histogram() +
  xlim(c(0,1)) +
  labs(x = "Percent Vaccinated", y="Count")
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
92040 ZIP code area is below and 92109 ZIP code area is above. 

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)
head(vax.36.all)

```
```{r}
ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with population above 36k are shown") +
  geom_hline(yintercept = 0.6645132, linetype= "dashed")
```

> Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

Would prefer to have time to get tested before meeting. 