---
title: 'Class 6: R Functions'
author: "Thisha Thiagarajan A15474979"
date: "10/14/2021"
output:
  html_document: default
  pdf_document: default
---

## A play with Rmarkdown

This is some plain text. I can make things **bold**. I can also make things *italic*.

```{r}
#This is a code chunk 
plot(1:10)
```

## R functions 

In today's class we are going to write a function together that grades some student work. 

Questions: 

> **Q1.** Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adequately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput”

```{r}
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)


mean(student1)

```

We need to drop the lowest grade. Try min() and which.min()

```{r}
min(student1)
which.min(student1)
```
Using which.min() creates a grade() function that works on vectors without NA elements. 

```{r}
grade <- function(x){
  low <- which.min(x)
  mean(x[-low])
} 

grade(student1)
grade(student2)
  
```

How can we account for vectors with NA elements. Try is.na(). Also try to read in online gradebook from csv file. I also used the apply function to assist in using the grade function on the csv file. 

```{r}
scores <- read.csv("https://tinyurl.com/gradeinput", row.names = 1)
grade <- function(x){
  #noticed csv file was being read as characters, giving NA output -> clarify
  #data type here to prevent that 
  x <-  as.numeric(x)
  #set all NA to 0 in grade vectors 
  x[is.na(x)] <- 0
  #find the position of the minimum value in the vector
  low <- which.min(x)
  #find the mean of all elements in the vector except the minimum value 
  mean(x[-low])
   
} 

grade(student1)
grade(student2)
grade(student3)
# apply allows the scores function to be applied to all rows of the df 
student_scores <- apply(scores, 1, grade)
student_scores
```
It works! 

> **Q2.** Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook?

```{r}
which.max(student_scores)
max(student_scores)
```


**Student 18** has the highest overall score of 94.5. 

> **Q3.** From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall?

I adapted the grade function to create a mean function that accounts for NA without dropping the lowest score. Using this I discovered that **homework 2** was the toughest on students (with an assignment average of 72.80). I used the apply function to apply my new meanNA function on the columns of the df scores. 

I used this method since it was more accurate than just using 
apply(scores, 2, mean, na.rm = TRUE) -> NA values are not counted as 0s here I believe.  

```{r}
meanNA <- function(x){
  #noticed csv file was being read as characters, giving NA output ->clarify
  #data type here to prevent that 
  x <-  as.numeric(x)
  #set all NA to 0 in grade vectors 
  x[is.na(x)] <- 0
  #find the mean of all elements in the vector except the minimum value 
  mean(x)
   
} 
apply(scores, 2, meanNA)
```

> **Q4.** Optional Extension: From your analysis of the gradebook, which homework was most predictive of overall score (i.e. highest correlation with average grade score)?

I think the cor() function will help with this. Using the cor() function, we see that **homework 5** has the highest correlation coefficient (closer to 1/-1, means higher correlation). 

```{r}

mask <- scores
mask[is.na(mask)] <- 0

#cor(student_scores, mask$hw1)
#cor(student_scores, mask$hw2)
#cor(student_scores, mask$hw3)
#cor(student_scores, mask$hw4)
#cor(student_scores, mask$hw5)

#run this more efficiently with the apply function 
apply(mask, 2, cor, student_scores)

```




> **Q5.** Make sure you save your Rmarkdown document and can click the “Knit” button to generate a PDF foramt report without errors. Finally, submit your PDF to gradescope.

Done

> Extension in Class07: Make it optional to drop lowest grade 

```{r}
grade2 <- function(x, drop.low = TRUE){
  #noticed csv file was being read as characters, giving NA output -> clarify
  #data type here to prevent that 
  x <-  as.numeric(x)
  #set all NA to 0 in grade vectors 
  x[is.na(x)] <- 0
  #find the position of the minimum value in the vector
  low <- which.min(x)
  #find mean, drop lowest score if needed using if/else statement
  if (drop.low){ 
    mean(x[-low])
  }
  else { 
    mean (x)
  }

}

grade2(student1, TRUE)
grade2(student1)
grade2(student1, FALSE)
grade(student1)
```



