---
title: "Class06HW"
author: "Thisha Thiagarajan A15474979"
date: "10/19/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

>**Section1** Improving Analysis Code by Writing Functions

> **a** Improve the code below. 

The function section1() simplifies the code. Now you can just call the function 
of df$column and it will perform  x <- (x - min(x)) / (max(x) - min(x)). 

```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)

df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a)) 
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b)) 
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c)) 
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))

df 

section1 <- function(x){
  x <- (x - min(x)) / (max(x) - min(x)) 
}
section1(df$a)
section1(df$b)
section1(df$c)
section1(df$d)

df
```

> **b** Improve the below example code for the analysis of protein drug interactions by abstracting the main activities in your own new function. 

> Q1. 

The read.pdb() function returns a dataframe object given input (PDB ID). 

> Q2. 

The trim.pdb() function selects a specific subset from the larger dataframe 
returned by read.pdf(). 

> Q3. 

The input parameter sse controls the marginal black and gray rectangles in the 
plots. Not including it would turn that feature off 
(ex. plotb3(s1.b, typ="l", ylab="Bfactor")  ). 
SSE represents secondary structure element and it basically annotates the 
positions of major secondary structure elements in the margins of the plot. 

> Q4. 

A better plot to compare different proteins would be to plot all 3 connected 
scatterplots on the same graph, making it easier to compare them to each other.
It would also be beneficial if we were able to obtain a quantifiable value for 
each of these scatterplots that allowed us to compare the proteins without 
relying on our subjective graphical analysis. 

> Q5. 

The dendrogram plot allows us to compare all three proteins to each other and
quantify the comparison,allowing us to see that protein 2 and protein 3 are more 
similar to each other in their Bfactor trends. 



```{r}
# Can you improve this analysis code? 
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug 
s2 <- read.pdb("1AKE") # kinase no drug 
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA") 
s2.chainA <- trim.pdb(s2, chain="A", elety="CA") 
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b 
s2.b <- s2.chainA$atom$b 
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor") 
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor") 
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

#Now create function to simplify

#documentation: 
#   input is the specification of file to be read (can be PDB ID)
#   the function creates a plot given a PDB ID to assist in analysis of protein 
#       drug interactions (compares residue to Bfactor)
#   output is the plot itself 
protein.drug <- function(x){ 
  #read.pdb
  s <- read.pdb(x)
  #trim.pdb
  s.chainA <- trim.pdb(s, chain="A", elety="CA")
  #set s.b
  s.b <- s.chainA$atom$b
  #plot
  plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor")
}

protein.drug("4AKE")
protein.drug("1AKE")
protein.drug("1E4Y")

#It works!

#Q5. creating dendrogram plot

hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```


```

