---
title: "class10"
author: "Thisha Thiagarajan A15474979"
date: "10/28/2021"
output: html_document
---

# Get the data 

Here we look at 538 Halloween candy data. We take the CSV file from their GitHub repo. 

```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)

```

> **Q1. How many different candy types are in this dataset?**

There are 85 different candy types in this data set. 

```{r}
nrow(candy)
```

> **Q2. How many fruity candy types are in the dataset?**

There are 38 candy types in the data set. 
```{r}
sum(candy$fruity)
```

> **Q3. What is your favorite candy in the dataset and what is it’s winpercent value?**

My favorite candy is the 3 Musketeers. It's winpercent value is 67.60294. 

```{r}
#row.names(candy)
candy["3 Musketeers", "winpercent"]
```

> **Q4. What is the winpercent value for “Kit Kat”?**

It's winpercent value is 76.7686. 

```{r}
candy["Kit Kat", "winpercent"]
```

> **Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?**

It's winpercent value is 49.6535. 

```{r}
candy["Tootsie Roll Snack Bars" , "winpercent"]
```

Use skimr to get a quick overview of the data set. 

```{r}
library("skimr")
skim(candy)
```

> **Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?**

Yes there is. It is wincpercent. The other columns look to be on a scale of 0-1, while this column is on a scale of 0-100. 

> **Q7. What do you think a zero and one represent for the candy$chocolate column?**

The 0 should represent that the specific candy does not have chocolate, while the 1 should represent that it does have chocolate. 

> **Q8. Plot a histogram of winpercent values**

```{r}
hist(candy$winpercent)
```
> **Q9. Is the distribution of winpercent values symmetrical?**

No it is skewed to the right. 

> **Q10. Is the center of the distribution above or below 50%?**

The center is below 50%. 

> **Q11. On average is chocolate candy higher or lower ranked than fruit candy?**

Chocolate has a higher mean than fruit (60.92 > 44.11). 
```{r}
#turn this into TRUE/FALSE to use 
inds <- as.logical(candy$chocolate)
#only pulls out true chocolates
choc <- candy[inds,]$winpercent
#find mean
mean(choc)

#repeat for fruit
inds2 <- as.logical(candy$fruity)
fru <- candy[inds2,]$winpercent
mean(fru)


```

> **Q12. Is this difference statistically significant?**

According to our t test, this difference is statistically significant. 

```{r}
#use t test to verify results 
t.test(choc, fru)
```


```{r}
head(candy[order(candy$winpercent),], n=5)

tail(candy[order(candy$winpercent),], n=5)

library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()

```

Improve this plot by reordering by the 'winpercent' value. 

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```
add color 

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent), col = my_cols) +
  geom_col() 
```

> **Q17. What is the worst ranked chocolate candy?**

Sixlets 

> **Q18. What is the best ranked fruity candy?**

Starburst

Plot pricepoint 

reeses miniatures 

```{r}

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text()

#improve labels with ggrepel

library(ggrepel)


# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```


```{r}
head(candy[order(candy$pricepercent),], n=5)

tail(candy[order(candy$pricepercent),], n=5)

```


```{r}

library(corrplot)
cij <- cor(candy)
corrplot(cij)
```


pca 

need to scale it first 

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)

plot(pca$x[,1:2])

```

